---
title: "Generating Random Portolios"
output: rmarkdown::github_document
---

```{r message = FALSE}
library(tidyverse)

source("utils.R")
```

One of the common problem in simulations is to generate random values that uniformly cover the search space.
This does not mean, however, that the solution is to simply draw from the uniform distribution.

In the `utils.R` file, I introduced a `generate_random_portfolio_weights()` function, which uses an approach suggested in a [StackExchange post](https://quant.stackexchange.com/a/49116), which in turn refers to an approach by [Rubinstein and Melamed (1998)](https://ie.technion.ac.il/~onn/Selected/AOR09.pdf). Values are effectively drawn from the exponential distribution and then normalized to the range between zero and one.

Let's see how this approach performs.
I draw 10000 portfolios, which _seems_ sufficient enough.
Each portfolio consists of three elements (there are three weights drawn), for the simplicity 

```{r}
n_portfolios <- 10000
n_weights <- 3
```

I generate the portfolios using the Rubinstein and Melamed approach (RM98 for short).

```{r}
random_portfolios_rm98 <- n_portfolios %>%
    replicate(generate_random_portfolio_weights(n_weights)) %>%
    t() %>%
    as_tibble() %>% 
    mutate(approach = "RM98")
```

As a benchmark I compare this with simply drawing three values from the uniform distribution, and scaling them to between zero and one.

```{r}
random_portfolios_unif <- n_portfolios %>%
    replicate(runif(n_weights)) %>%
    apply(2, function(row) row / sum(row)) %>% 
    t() %>%
    as_tibble() %>% 
    mutate(approach = "Uniform")
```

A purely visual comparison reveals that the RM98 approach _looks_ uniformly spread, as opposed to the benchmark approach.

```{r}
random_portfolios <- rbind(random_portfolios_rm98, random_portfolios_unif)

random_portfolios %>%
    ggplot(aes(x = V1, y = V2)) +
    geom_point(alpha = 0.1, shape = 16, size = 2) +
    facet_grid(. ~ approach)
```

Of course, to mathematically prove that the RM98 is indeed what we want is a completely different story.
This, however, is beyond the scope of this repository, at least for now.
