---
title: "Portfolio Optimisation for NN Investment Parners"
output: rmarkdown::github_document
---

```{r message = FALSE}
library(tidyverse)
library(tidyquant)
library(timetk)
library(lubridate)

source("utils.R")
```

```{r}
HISTORICAL_DATA_PATH <- "./historical-fund-data"
METADATA_PATH <- "./fund-metadata.csv"
```

```{r}
START_DAY <- ymd("2005-01-01")
END_DAY <- ymd("2020-12-31")
```

```{r, message = FALSE}
funds_metadata <- METADATA_PATH %>% 
    read_csv()

funds_data <- HISTORICAL_DATA_PATH %>% 
    list_nn_funds() %>% 
    bind_rows() %>% 
    left_join(funds_metadata) %>% 
    select(-fund) %>% 
    rename(fund = fund_short_name)
```

These are all the funds used in the analysis.

```{r}
ggplot(funds_data, aes(x = date, y = value, color = fund)) +
    geom_line() +
    labs(
        title = "NN Funds",
        subtitle = "All Available Funds",
        caption = "Source: NN Investment Partners, michaltkaczyk's estimations",
        color = "Fund") +
    xlab("Time") +
    ylab("Index")
```

Let's do some data processing here.

First of all, let's drop all founds that are not fully contained within a specified time period.

```{r}
funds_old_enough <- funds_data %>%
    group_by(fund) %>%
    summarise(start_day = min(date)) %>%
    filter(start_day <= START_DAY) %>% 
    pull(fund)

funds_data <- funds_data %>%
    filter(fund %in% funds_old_enough & date >= START_DAY & date <= END_DAY)
```

Now, let's change the format to `xts`, for convenience of calculations later on.
In the meantime, I am replacing all `NA` values with zeros, so that they are not dropped in the next step.

```{r}
funds_data <- funds_data %>%
    pivot_wider(names_from = fund, values_from = value) %>% 
    tk_xts()
```

The next step is to change the granularity from daily to monthly.
This should be enough for my purposes.

```{r}
funds_data <- funds_data %>% 
    to.monthly(indexAt = "yearmon", OHLC = FALSE)
```

Finally let's change the absolute values to monthly rates of change. 

```{r}
funds_data <- funds_data %>% 
    ROC()
```

Let's see where we got after all of those manipulations.

```{r}
funds_data_dates <- funds_data %>% 
    index() %>% 
    as_tibble() %>% 
    rename(date = value)

funds_data_long <- funds_data %>%
    as_tibble() %>%
    add_column(funds_data_dates) %>% 
    pivot_longer(!date, "fund") %>%
    replace_na(list(value = 0)) %>% 
    group_by(fund) %>% 
    arrange(date) %>% 
    mutate(value = cumprod(value + 1)) %>% 
    ungroup()

ggplot(funds_data_long, aes(x = date, y = value, color = fund)) +
    geom_line() +
    labs(
        title = "NN Funds",
        subtitle = "All Available Funds",
        caption = "Source: NN Investment Partners, michaltkaczyk's estimations",
        color = "Fund") +
    xlab("Time") +
    ylab("Index")
```

Let's check some manually created portfolios' performance.

```{r, message = FALSE, warning = FALSE}
manual_portfolios <- rbind(
    c(1, 0, 0, 0, 0),
    c(1, 1, 0, 0, 0),
    c(1, 0, 1, 0, 0),
    c(1, 0, 0, 1, 0),
    c(1, 0, 0, 0, 1),
    c(1, 1, 1, 1, 1)) %>% 
    apply(1, function(x) x / sum(x)) %>% 
    t()

manual_portfolios_results <- list()

for (portfolio in 1:NROW(manual_portfolios)) {
    manual_portfolios_results[[portfolio]] <- 
        Return.portfolio(funds_data, manual_portfolios[portfolio, ]) %>%
        as_tibble() %>% 
        add_column(date = index(funds_data)) %>% 
        mutate(portfolio = paste("portfolio", portfolio)) %>% 
        rename(value = portfolio.returns)
}

portfolio_comparison <- manual_portfolios_results %>%
    bind_rows() %>% 
    group_by(portfolio) %>% 
    mutate(value = cumprod(value + 1) * 100)
```

```{r}
ggplot(portfolio_comparison, aes(x = date, y = value, color = portfolio)) +
    geom_line() +
    labs(
        title = "Performance of Manually Selected NN Portfolios",
        subtitle = paste("Between", START_DAY, "and", END_DAY),
        caption = "Source: NN Investment Partners, michaltkaczyk's estimations",
        color = "Portfolio") +
    xlab("Time") +
    ylab("Index")
```

What if we tried randomly generating the portfolios?

```{r}
n_portfolios <- 100

random_portfolios <-
    n_portfolios %>%
    replicate(generate_random_portfolio_weights(5)) %>% 
    t()

random_portfolios_results <- list()

for (portfolio in 1:NROW(random_portfolios)) {
    random_portfolios_results[[portfolio]] <- 
        Return.portfolio(funds_data, random_portfolios[portfolio, ]) %>%
        as_tibble() %>% 
        add_column(date = index(funds_data)) %>% 
        mutate(portfolio = paste("portfolio", portfolio)) %>% 
        rename(value = portfolio.returns)
}

random_portfolio_comparison <- random_portfolios_results %>%
    bind_rows() %>% 
    group_by(portfolio) %>% 
    mutate(value = cumprod(value + 1) * 100)
```

```{r}
ggplot(random_portfolio_comparison, aes(x = date, y = value, group = portfolio)) +
    geom_line(alpha = 0.05) +
    labs(
        title = "Performance of Randomly Selected NN Portfolios",
        subtitle = paste("Between", START_DAY, "and", END_DAY),
        caption = "Source: NN Investment Partners, michaltkaczyk's estimations",
        color = "Portfolio") +
    xlab("Time") +
    ylab("Index")
```

```{r}
funds_data_cov <- funds_data %>% cov(use = "na.or.complete")
funds_data_avg <- funds_data %>% apply(2, mean, na.rm = TRUE)

random_portfolios_results_avg <- matrix(nrow = NROW(random_portfolios), ncol = 2)
colnames(random_portfolios_results_avg) <- c("return", "risk")

for (portfolio in 1:NROW(random_portfolios)){
    random_portfolios_results_avg[portfolio, 1] <- sum(random_portfolios[portfolio, ] * funds_data_avg) * sqrt(12) * 100
    random_portfolios_results_avg[portfolio, 2] <- as.numeric(sqrt(random_portfolios[portfolio, ] %*% funds_data_cov %*% random_portfolios[portfolio, ])) * 12 * 100
}

random_portfolios_results_avg <-
    random_portfolios_results_avg %>%
    as_tibble() %>% 
    bind_cols(random_portfolios)

ggplot(random_portfolios_results_avg, aes(x = risk, y = return)) +
    geom_point() +
    labs(
        title = "Average Performance of Randomly Selected NN Portfolios",
        caption = "Source: NN Investment Partners, michaltkaczyk's estimations") +
    xlab("Risk") +
    ylab("Return")
```
